---
title: "codebook.Rmd"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(dplyr)
```

## Codebook

Codebook for assignment: "Getting and Cleaning Data Course Project"

## Data source

As described in the instructions the code was downloaded from https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip
```{r}
    if(!file.exists("data/dataset.zip")){ 
        flog.info("downloading dataset")
        fileUrl = "https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip"
        download.file(fileUrl,"data/dataset.zip",method="curl") 
    }
```

## Transformations

### Stage 1
Read both the test and training files for activity.
```{r}
    test_activity <- read.table("data/UCI HAR Dataset/test/Y_test.txt")
    train_activity <- read.table("data/UCI HAR Dataset/train/Y_train.txt")
```
Combine into a single table.
```{r}
    all_activities <- rbind(test_activity,train_activity)
    names(all_activities)<-c("activity")
```
Convert activities into readable labels. See below for sample output
```{r}
    activity_labels <- read.table("data/UCI HAR Dataset/activity_labels.txt")
    all_activities$activity <- factor(all_activities$activity, levels=activity_labels[,1], labels = activity_labels[,2])
    sample_n(all_activities,4)
```

### Stage 2
Read both the test and training subjects and combine them into a single table. See below for sample output.
```{r}
    test_subjects <- read.table("data/UCI HAR Dataset/test/subject_test.txt")
    train_subjects <- read.table("data/UCI HAR Dataset/train/subject_train.txt")
    subjects<-rbind(test_subjects,train_subjects)
    names(subjects)<-c("subject")
    sample_n(subjects,4)
```

### Stage 3
Read both test and training features and combine then into a single table.
```{r}
    test_features <- read.table("data/UCI HAR Dataset/test/X_test.txt")
    train_features <- read.table("data/UCI HAR Dataset/train/X_train.txt")
    features_names<-fread("data/UCI HAR Dataset/features.txt")
    features<-rbind(test_features,train_features)
```

### Stage 4
From the output of stage 3 only retain features that have the work mean or std in it.
```{r}
    names(features)<-features_names$V2
    features<-subset(features, select=grep(x=colnames(features), "mean\\(\\)|std\\(\\)"))
```

### Stage 5
Build a data frame called Data that combines subjects, activities and features.
```{r}
    Data<-cbind(cbind(subjects,all_activities),features)
    sample_n(Data,2)
```

### Stage 6
Transform feature names into more descriptive names.
```{r}
    names(Data)<-gsub("^t", "time", names(Data))
    names(Data)<-gsub("^f", "frequency", names(Data))
    names(Data)<-gsub("Acc", "Accelerometer", names(Data))
    names(Data)<-gsub("Gyro", "Gyroscope", names(Data))
    names(Data)<-gsub("Mag", "Magnitude", names(Data))
    names(Data)<-gsub("BodyBody", "Body", names(Data))
    names(Data)<-gsub("-std\\(\\)", "StandardDeviation", names(Data))
    names(Data)<-gsub("-mean\\(\\)", "Mean", names(Data))
```

### Stage 6
Group by subject and activity, summarize the features.
```{r}
    Data2<-Data %>% group_by(subject,activity) %>% summarise_each(funs(mean))
```
